---
title: "pan-NF multiPLIER analysis for Landscape Paper"
author: "Robert Allaway"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Setup

```{r setup}
library(tximport)
library(tidyverse)
library(synapser)
library(org.Hs.eg.db)
library(PLIER)
```

## Login to Synapse

```{r include=FALSE}
synLogin()
```

## Get annotation maps and multiPLIER code

```{r}
source('https://raw.githubusercontent.com/greenelab/multi-plier/7c56a2867f9f08f7bcd3617875c7c2dbe886cbeb/util/plier_util.R')

plier_model <- readr::read_rds(synGet("syn18689545")$path) ## Provided by the CCDL (Jaclyn Taroni, Casey Greene, et al)

`%>%` <- magrittr::`%>%`

```

## Get expression data

```{r message=FALSE, warning=FALSE}
tx2gene_df <- synGet('syn18482848')$path %>% readr::read_tsv()


metadata <- synTableQuery("SELECT * FROM syn21046851", includeRowIdAndRowVersion=F)$asDataFrame() #this is a frozen table of ("SELECT * FROM syn16858331 where fileFormat = 'sf' and accessType in ('PUBLIC','REQUEST ACCESS') and isCellLine = 'FALSE'") as well as ("SELECT * FROM syn20629788 where fileFormat = 'sf' and tumorType = 'plexiform neurofibroma'")


metadata$isCellLine <- toupper(metadata$isCellLine)

rownames(metadata) <- metadata$id

metadata[is.na(metadata)] <- "NA"

salmon_entities <- sapply(metadata$id, synGet)

salmon_filepaths <- sapply(salmon_entities, function(x){
  x$path
})

tx_data <- tximport(salmon_filepaths, type = 'salmon', tx2gene = tx2gene_df,
                          countsFromAbundance = "no", ignoreTxVersion = TRUE)
```


## Map IDs 

Map Ensembl ids to HUGO identifiers for multiPLIER. Get rid of duplicated genes and filter out low abundance genes (sum of gene counts < 10 removed).
```{r}
abundance <- tx_data$abundance %>% as.data.frame() 

mapped_list <- mapIds(org.Hs.eg.db, keys = rownames(abundance), column = "SYMBOL", 
                      keytype = "ENSEMBL", multiVals = "list")

mapped_df <- reshape2::melt(mapped_list)

colnames(mapped_df) <- c("SYMBOL", "ENSEMBL")

abundance$gene_id <- rownames(abundance)

# let's use the mappings to reannotate our data
abundance <- mapped_df %>%
  # removing anything in the SYMBOL column that is NA
  filter(!is.na(SYMBOL)) %>%
  # join by Ensembl gene ids, retaining only Ensembl gene ids in both
  inner_join(y = abundance, by = c("ENSEMBL" = "gene_id")) %>% 
  dplyr::select(-ENSEMBL) %>% 
  add_count(SYMBOL)

duplicate_genes <- abundance %>% filter(n > 1) %>% dplyr::select(-n)

abundance_filt <- abundance  %>% 
  filter(n == 1) %>% 
  tibble::column_to_rownames("SYMBOL") %>% 
  dplyr::select(-n)  %>%
  as.matrix()

abundance_filt_cutoff <- abundance_filt[rowSums(abundance_filt) > 10,]

```

# multiPLIER

## Filtering out LVs

```{r}
mp_res <- GetNewDataB(abundance_filt_cutoff, plier_model)
```

### Filter by standard deviation 

Most of the latent variables have low standard deviation across this cohort of samples, and thus are probably not interesting for comparing across or within tumor groups. A cutoff of SD > 0.075 seems to be a reasonable cutoff to filter out the least variable LVs. 

```{r}
mp_res_tidy <- mp_res %>% 
  as.data.frame(.) %>% 
  tibble::rownames_to_column('latent_var') %>% 
  tidyr::gather("id","value",-latent_var) %>% 
  group_by(latent_var) %>% 
  mutate(sd_value = sd(value)) %>% 
  ungroup() %>% 
  mutate(trunc_label = stringr::str_trunc(latent_var, 10, "right")) %>% 
  left_join(metadata)

ggplot(data= mp_res_tidy %>% dplyr::select(latent_var,sd_value) %>% distinct()) +
  geom_histogram(aes(x=sd_value))

ggplot(data = mp_res_tidy) +
  geom_boxplot(aes(x=reorder(trunc_label, -sd_value), y = value, color = sd_value > 0.075))

```

It is probably a safe assumption that (at least for subtyping these tumors), the latent variables with high variance across the cohort are the most "interesting." We won't filter these out a priori because we have some downstream analyses that might use te low sd data as well. 

```{r eval=FALSE, include=FALSE}

# mp_res_tidy_filt <- mp_res_tidy %>% 
#   filter(sd_value > 0.05)
# 
# mp_res_filt <- mp_res[rownames(mp_res) %in% unique(mp_res_tidy_filt$latent_var),]

```

### Filter by Correlation

We'll correlate the loadings of the latent variables to see if any have huge overlap between one another. In the plot below, if "high SD" is true, we are looking at the correlation of a latent variable where the standard deviation is 0.075 to another LV with a SD > 0.075. All other comparisions are in red. 

```{r}

plier_loadings <- plier_model$Z 
colnames(plier_loadings) <- rownames(mp_res)

plier_cor <- cor(plier_loadings) 

interesting_lvs <- mp_res_tidy$latent_var 

plier_cor_tidy <- plier_cor %>% 
  as.data.frame() %>% 
  tibble::rownames_to_column('lv1') %>% 
  tidyr::gather("lv2","cor",-lv1) %>% 
  filter(lv1 != lv2) %>% 
  distinct() %>%
  mutate(high_sd=case_when((lv1 %in% interesting_lvs & lv2 %in% interesting_lvs) ~ TRUE,
                               (!lv1 %in% interesting_lvs | !lv2 %in% interesting_lvs) ~ FALSE))

ggplot(data = plier_cor_tidy) + 
  geom_density(aes(x = cor, color = high_sd))

```

And toss out latent variables with a Pearson correlation >0.5 to other latent variables.

```{r}
high_cor_rm <- plier_cor_tidy %>% filter(cor > 0.5 & high_sd == TRUE) %>% group_by(cor) %>% dplyr::slice(2) %>% ungroup()

mp_res <- mp_res[!rownames(mp_res) %in% high_cor_rm$lv1,]

mp_res_tidy_filt <- filter(mp_res_tidy, !latent_var %in% high_cor_rm$lv1)
```

```{r eval=FALSE, include=FALSE}
##upload results

mp_res_tidy_final <- dplyr::select(mp_res_tidy_filt, id, name, latent_var, value, projectId, consortium, diagnosis, tumorType, fundingAgency, individualID, nf1Genotype, nf2Genotype, species, isCellLine, studyId, studyName, specimenID, sex, cellType, modelOf, experimentalCondition)

# cols <- list(
#   Column(name = "id", columnType = "ENTITYID"),
#   Column(name = "name", columnType = "STRING", maximumSize = 100),
#     Column(name = "latent_var", columnType = "LARGETEXT"),
#     Column(name = "value", columnType = "DOUBLE"),
#     Column(name = "projectId", columnType = "ENTITYID"),
#     Column(name = "consortium", columnType = "STRING", maximumSize = 100),
#     Column(name = "diagnosis", columnType = "STRING", maximumSize = 100),
#     Column(name = "tumorType", columnType = "STRING", maximumSize = 100),
#     Column(name = "fundingAgency", columnType = "STRING", maximumSize = 100),
#     Column(name = "individualID", columnType = "STRING", maximumSize = 100),
#     Column(name = "nf1Genotype", columnType = "STRING", maximumSize = 100),
#     Column(name = "nf2Genotype", columnType = "STRING", maximumSize = 100),
#     Column(name = "species", columnType = "STRING", maximumSize = 100),
#     Column(name = "isCellLine", columnType = "STRING", maximumSize = 100),
#       Column(name = "studyId", columnType = "STRING", maximumSize = 100),
#       Column(name = "studyName", columnType = "STRING", maximumSize = 100),
#       Column(name = "specimenID", columnType = "STRING", maximumSize = 100),
#       Column(name = "sex", columnType = "STRING", maximumSize = 100),
#       Column(name = "cellType", columnType = "STRING", maximumSize = 100),
#       Column(name = "modelOf", columnType = "STRING", maximumSize = 100),
#       Column(name = "experimentalCondition", columnType = "STRING", maximumSize = 100))
# 
# schema <- Schema(name = "Pan-NF MultiPLIER Results for Landscape Paper", parent = "syn21046734", columns = cols)
# table <- Table(schema, mp_res_tidy_final)
# table <- synStore(table)

#Table and schema exist from first time code was run, so for re-runs we can just update it by: 

res <- synTableQuery("SELECT * FROM syn21046991")
del <- synDelete(res)

tab <- Table('syn21046991', mp_res_tidy_final)
tab <- synStore(tab)

##upload loading data - commented out so we don't redo this 
# table_2 <- synBuildTable("MultiPLIER LV Loading", parent = "syn21046734", plier_cor_tidy)
# synStore(table_2)

```




```{r}
sessionInfo()
```

