---
title: "Run MetaViper on NF1 Samples"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()

```

## First get all gene expression data
Gene expression data is spread across four synapse projects:
- the CNF dataset
- the JHU Biobank
- the CBTTC dataset
- the WashU biobank

Since permissions are dispersed, we need to collect all the data. 

```{r table, warning=FALSE}
tabids<-synTableQuery('select distinct tableId from syn21221980')$asDataFrame()

vars="specimenID,individualID,Symbol,totalCounts,zScore,tumorType,nf1Genotype,sex"
full.tab<-do.call(rbind,lapply(tabids$tableId,function(x) synTableQuery(paste('select',vars,'from',x))$asDataFrame()))



```

## Run Metaviper

To run metaviper we need to load all the aracne networks and run one-by-one. This requires a bit of prep. First we have to load the aracne networks, and map the table to entrez identifiers. 

```{r viper prep, echo=FALSE, warning=FALSE}
#get aracne networks
 require(aracne.networks)
                #get aracne networks
net.names <- data(package="aracne.networks")$results[, "Item"]
all.networks <- lapply(net.names,function(x) get(x))
names(all.networks) <- net.names

#now map the gene identifiers
library(AnnotationDbi)
library(org.Hs.eg.db)
x <- org.Hs.egSYMBOL2EG
mapped_genes <- AnnotationDbi::mappedkeys(x)
xx <- AnnotationDbi::as.list(x[mapped_genes])

genes=unique(full.tab$Symbol)
inds=match(genes,names(xx))
genes<-genes[which(!is.na(inds))]
inds<-inds[which(!is.na(inds))]
entrez_list=data.frame(gene=genes,entrezgene=unlist(sapply(xx[inds],function(y)y[1])))
```

Then we can actually run the tool!
```{r run viper, echo=FALSE, warning=FALSE}
#now create the data matrix
require(dplyr)
tidied.df<-full.tab%>%rename(gene='Symbol')%>%left_join(entrez_list,by='gene')

#tidied.df$specimenID<-make.names(tidied.df$specimenID)

combined.mat<-reshape2::acast(tidied.df,entrezgene~specimenID,value.var="totalCounts",fun.aggregate=function(x) mean(x,na.rm=T))

#some genes are only quantified in some samples
missing.genes<-unique(which(is.na(combined.mat),arr.ind=T)[,1])

if(length(missing.genes)>0)
  combined.mat<-combined.mat[-missing.genes,]

library(viper)
res <- viper::viper(combined.mat,all.networks)

rdf<-data.frame(res,entrezgene=rownames(res),check.names=F)
tdf<-tidyr::gather(rdf,key=specimenID,value=metaviperscore,-entrezgene)

gn<-tdf%>%left_join(entrez_list,by='entrezgene')
pats<-tidied.df%>%select(specimenID,individualID,tumorType,nf1Genotype,sex)%>%distinct()


fdf<-gn%>%left_join(pats,by='specimenID')
```

Now to view results
```{r results}
DT::datatable(head(fdf))
```

Once it's in a data frame we can store back on the Synapse project.

```{r store on synapse,echo=FALSE}

tab<-synBuildTable('Metaviper NF1 Protein scores',parent='syn21046734',fdf)
synStore(tab)