---
title: "02-Immune Sex Differences"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()
require(tidyverse)
```

## Introduction
This is a first stab at checking immune annotation predictions to see if the presence/absence of immune cells is any different between sexes.

First get the tumor de-convolution results for NF1 tumors from synapse, combine non-standard tumor types and handle basic renaming of variables.
```{r,warning=FALSE}
tab<-synapser::synTableQuery("select * from syn20710536")$asDataFrame()

tab<-subset(tab,diagnosis=='Neurofibromatosis 1')

tab$tumorType[which(tab$tumorType%in%c("Ependymoma","Ganglioglioma","Low Grade Glioma","High Grade Glioma"))]<-'Other'

tab$tumorType[which(tab$tumorType=="Malignant peripheral nerve sheath tumor")]<-"Malignant Peripheral Nerve Sheath Tumor"

tab<-subset(tab,tumorType!='Other')
```

## Check tumor-specific sex differences
```{r}
##now what do we see on a tissue level? 
res.c<-tab%>%subset(method!='xcell')%>%
  spread(key=sex,value=score)%>%
  group_by(method,tumorType,cell_type)%>%
  mutate(pval=t.test(female,male)$p.value)%>%
  select(method,cell_type,pval,tumorType)%>%distinct()%>%
  group_by(method)%>%
  mutate(correctedP=p.adjust(pval))

sigs<-subset(res.c,pval<0.05)

DT::datatable(sigs)
```

There are numerous significant cell-type/tumor combinations.
```{r,warning=FALSE}
library(ggpubr)

for(ct in unique(sigs$method)){
  sigs.t=subset(sigs,method==ct)
  for(tu in unique(sigs.t$tumorType)){
    sigs.tt=subset(sigs.t,tumorType==tu)
    tab.t=subset(tab,method=ct)%>%subset(tumorType==tu)%>%subset(cell_type%in%sigs.tt$cell_type)

  p<-ggboxplot(tab.t,x='sex',y='score',facet.by='cell_type',color='sex',palette='jco')+stat_compare_means(method='t.test')+ggtitle(paste(ct,tu,' significant differences'))
  print(p)
  
  #  tab.p<-subset(tab.t,method==meth)%>%subset(cell_type%in%(sigs.t$cell_type))
# p<-ggplot(tab.t,palette='jco')+geom_boxplot(aes(x=cell_type,fill=sex,y=score))+facet_grid(.~tumorType)+ theme(axis.text.x = element_text(angle = 90, hjust = 1))
 # p<-ggboxplot(subset(tab.t,cell_type=='Neutrophil'),x='tumorType',y='score',color='sex',palette='jco')+stat_compare_means(method='t.test')+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
 # print(p)
}
  
}

```
There are some interesting tumor-specific differences, primarily in cNFs, but also some in MPNSTs (with their lone male)

```{r,warning=FALSE}
res<-tab%>%spread(key=sex,value=score)%>%
  group_by(method,cell_type)%>%
    mutate(pval=t.test(female,male)$p.value)%>%
  select(method,cell_type,pval)%>%distinct()%>%
  group_by(method)%>%
  mutate(correctedP=p.adjust(pval))


sigs<-subset(res,pval<0.05)
sigs
```
## Pan-tumor sex differences

```{r}

for(meth in unique(tab$method)){
  tab.p<-subset(tab,method==meth)%>%subset(cell_type%in%(sigs$cell_type))

  p<-ggboxplot(tab.p,x='sex',y='score',facet.by='cell_type',color='sex',palette='jco')+stat_compare_means(method='t.test')+ggtitle(paste(meth,'significant differences'))
  print(p)
}

```  

