---
title: "06 - Cutaneous Neurofibroma patient Variables"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()
require(tidyverse)
```

## Synapse tables with data in them
We have two tables on synapse that have cNF patient data in them
- [Patient Table](https://www.synapse.org/#!Synapse:syn7342635/tables/)
- [Sample Table](https://www.synapse.org/#!Synapse:syn5556216/tables/)


```{r,warning=FALSE}
pat.tab<-synapser::synTableQuery("SELECT Patient,Race,Gender,Age,Pain,Itching FROM syn7342635")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))

samp.tab<-synapser::synTableQuery("SELECT Patient,TumorNumber,Length_in_mm FROM syn5556216")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))
pat.tab

samp.tab
```
For the purposes of this analysis we want to have just the age, sex, tumor size and and 'reformed' patient name
```{r, warning=FALSE}
full.tab<-samp.tab%>%left_join(pat.tab,by='Patient')%>%
  mutate(specimenID=paste0('patient',Patient,'tumor',TumorNumber))
full.tab
```

Now we can evaluate other things

## Plot the data
```{r}
##now what do we see on a tissue level? 
require(ggplot2)

p<-ggplot(full.tab)+geom_point(aes(x=Age,y=Length_in_mm,color=Itching,shape=Gender))+ggtitle("Age by tumor size with itching")

print(p)

p<-ggplot(full.tab)+geom_point(aes(x=Age,y=Length_in_mm,color=Itching,shape=Pain))+ggtitle("Age by tumor size with pain")
print(p)
```
So the pain/itching variables only go by patient, so we can't figure out which samples are in pain/itching.  


## Get the expression data and see what correlates
```{r}
exp.data<-synTableQuery('select * from syn20449214 where tumorType=\'Cutaneous Neurofibroma\'')$asDataFrame()

data.with.var<-exp.data%>%left_join(full.tab,by='specimenID')

#now compute the correlation with size for each transcript...?
gene.cors=data.with.var%>%group_by(Symbol)%>%mutate(corVal=cor(zScore,Length_in_mm))

top.genes=select(gene.cors,corVal)%>%distinct()%>%arrange(desc(corVal))%>%select(Symbol)

bottom.genes=select(gene.cors,corVal)%>%distinct()%>%arrange(corVal)%>%select(Symbol)
 
```

Now we have the genes most correlated with tumor size
```{r}
top.genes
```
and the genes least correlated
```{r}
bottom.genes
```

Now what do these look like?

```{r}

ggplot(subset(data.with.var,Symbol%in%c(top.genes$Symbol[1:10])))+geom_point(aes(x=Length_in_mm,y=zScore,col=Symbol,shape=sex))+ggtitle('10 most correlated genes')

ggplot(subset(data.with.var,Symbol%in%c(bottom.genes$Symbol[1:10])))+geom_point(aes(x=Length_in_mm,y=zScore,col=Symbol,shape=sex))+ggtitle('10 most anti-correlated genes')

```

We should do pathway enrichment of these genes, see if they are doing anything. The fact that HOXB3 is there is already interesting. What else could there be???

## Next steps
I think this is actually pretty cool - we can try to identify which are the 'driving' genes of cNF growth. 
* What pathways are enriched in these genes?
* Are these correlations statistically significant?
* Are there differences in correlated genes b/w male and female patients?
* What are these genes doing in pNFs and MPNSTs? 
* Are any of these correlated with immune reponse (conversely are immune signatures correlated with this size variable)?

Are there other questions we can answer?
