---
title: "07-Tumor Deconvolution-Latent Variable Correlation"
author: "Robert Allaway"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##compare tumor immune scores to lv expression

library(synapser)
synLogin()
library(tidyverse)

deconv_scores='syn20710536'
```

### Immune predictions correlated with multiplier latent variable expression

First we get the LV multiplier values from the nf-lv-viz repo and deconvolution predictions from Synapse

```{r, warning=FALSE}
#get immune predictions
dtab<-synapser::synTableQuery(paste('select * from',deconv_scores))$asDataFrame()%>%
  rename(immScore='score')

##get latent variable scores
mp_res<-synTableQuery("SELECT * FROM syn21046991")$filepath %>% 
  readr::read_csv() %>% 
  filter(!is.na(tumorType),
         !grepl('xenograft', specimenID, ignore.case = T),
         !specimenID %in% c("BI386-004","CW225-001","DW356-002",
                            "JK368-003", "SK436-005"))

dtab<-subset(dtab,method!='xcell')

combined=dtab%>%select(c(cell_type,method,specimenID,immScore))%>%
  left_join(mp_res,by='specimenID')
```

Then we can compute the correlation of each cell type measurement with each protein
```{r, warning=FALSE}
corVals=combined%>%group_by(cell_type,latent_var,method)%>%summarize(corVal=cor(immScore,value,method = "spearman"))

DT::datatable(corVals)
```

### Plotting LV correlations
We are interested in particular protein signaling activity that might give rise to specific immune phonetypes.

```{r, warning=FALSE}
##now how do we bracket them?
##plot correlation distributions by cell type and method. 
require(ggplot2)
p<-ggplot(corVals)+geom_boxplot(aes(x=cell_type,y=corVal,fill=method))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ggtitle("Correlation of MultiPLIER Latent Variables with predicted cell type")
print(p)
```

There are some latent variables that show up as highly correlated. By choosing a threshold, we can evaluate what they are in more detail.

```{r, warning=FALSE}
corthresh=0.6

##now filter to the cell types with correlated latent variables (or anticorrelated)
cor_cell_types=subset(corVals,abs(corVal)>corthresh)%>%ungroup()%>%
  select(cell_type,method)%>%unique()
print(paste('we found',nrow(cor_cell_types),'cell types with some latent variable correlation greater than',corthresh))


apply(cor_cell_types,1,function(x){
  ct=x[['cell_type']]
  m=x[['method']]

  #for each gene and cell type
  lvs=subset(corVals,cell_type==ct)%>%
        subset(abs(corVal)>corthresh)%>%
    subset(method==m)%>%arrange(desc(corVal))%>%
      ungroup()

    if(nrow(lvs)>12){
    new.corthresh=format(lvs$corVal[15],digits=3)
    lvs=lvs[1:12,]
  }else{
    new.corthresh=corthresh
  }

  scores=subset(combined,latent_var%in%lvs$latent_var)%>%subset(cell_type==ct)%>%subset(method==m)

  p2<- ggplot(scores %>% mutate(latent_var = stringr::str_trunc(latent_var, 30)))+
      geom_point(aes(x=immScore,y=value,
          col=latent_var,shape=tumorType))+
    scale_x_log10()+
      ggtitle(paste(m,'predictions of',ct,'correlation >',new.corthresh))
  print(p2)
 # ggsave(paste0(m,'predictions of',gsub(" ","",gsub("/","",ct)),'cor',new.corthresh,'.pdf'))
})

#parentid='syn20710537'
#for(fi in list.files('.')[grep('tions',list.files('.'))])
#  synapser::synStore(synapser::File(fi,parentId=parentid,annotations=list(resourceType='analysis',isMultiSpecimen='TRUE',isMultiIndividual='TRUE')),used=c(deconv_scores,metaviper_scores),executed=this.script)
```

