---
title: "08 - Cutaneous Neurofibroma Tumor Growht and Immune Cor"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()
require(tidyverse)
```

## Synapse tables with data 
We have two tables on synapse that have cNF patient data in them
- [Patient Table](https://www.synapse.org/#!Synapse:syn7342635/tables/)
- [Sample Table](https://www.synapse.org/#!Synapse:syn5556216/tables/)

We also have Immune Data
- [Tumor Deconvolution Table](https://www.synapse.org/#!Synapse:syn20710536/tables/)

Let's see if there are any correlations between immune populations and tumor growth.


```{r,warning=FALSE}
pat.tab<-synapser::synTableQuery("SELECT Patient,Race,Gender,Age,Pain,Itching FROM syn7342635")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))

samp.tab<-synapser::synTableQuery("SELECT Patient,TumorNumber,Length_in_mm FROM syn5556216")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))


```

## Merge data together
For the purposes of this analysis we want to have just the age, sex, tumor size and and 'reformed' patient name
```{r, warning=FALSE}
full.tab<-samp.tab%>%left_join(pat.tab,by='Patient')%>%
  mutate(specimenID=paste0('patient',Patient,'tumor',TumorNumber))
DT::datatable(full.tab)
```

Now we can get the immune data!

## Collect Immune Data

Now we can download the imune data and evaluate any correlations betwen scores and tumor size
```{r,warning=FALSE}
imm.data<-synTableQuery("SELECT * FROM syn20710536 where tumorType='Cutaneous Neurofibroma'")$asDataFrame()
data.with.var<-imm.data%>%left_join(full.tab,by='specimenID')

#now coampute the correlation with size for each transcript...?
tum.cors=data.with.var%>%group_by(method,cell_type)%>%
  mutate(corVal=cor(score,Length_in_mm))%>%
  arrange(desc(abs(corVal)))%>%
  select(cell_type,method,corVal)%>%distinct()

DT::datatable(tum.cors)
```

So this is interesting - not all cell types are associated with tumor growth, specifically Mast cells are not. NK cells, macrophages, B cells - they are. Yes the correlation values are as high as we'd like, but it's a start.

```{r,warning=FALSE}

cs.cells=subset(tum.cors,abs(corVal)>0.35)%>%
    subset(method=='cibersort')%>%ungroup()%>%
    select(cell_type)%>%unlist()

subset(data.with.var,method=='cibersort')%>%
    subset(cell_type%in%cs.cells)%>%
    ggplot()+geom_point(aes(x=score,y=Length_in_mm,col=sex))+facet_grid(cell_type~.)

mc.cells=subset(tum.cors,abs(corVal)>0.35)%>%
    subset(method=='mcp_counter')%>%ungroup()%>%
    select(cell_type)%>%unlist()

subset(data.with.var,method=='mcp_counter')%>%
    subset(cell_type%in%mc.cells)%>%
    ggplot()+geom_point(aes(x=score,y=Length_in_mm,col=sex))+facet_grid(cell_type~.)
```

So it seems like the dendritic cells, monocytes, and macrophages seem to have a weak role in tumor growth. 

## Any correlation with itching

Perhaps itching is driving this?
```{r,warning=FALSE}

subset(data.with.var,method=='cibersort')%>%
    subset(cell_type%in%cs.cells)%>%
    ggplot()+geom_point(aes(x=score,y=Length_in_mm,col=Itching))+facet_grid(cell_type~.)


subset(data.with.var,method=='mcp_counter')%>%
    subset(cell_type%in%mc.cells)%>%
    ggplot()+geom_point(aes(x=score,y=Length_in_mm,col=Itching))+facet_grid(cell_type~.)

```

Doesn't seem to be. But we can do a test to see if Itching is correlated
```{r, warning=FALSE}
data.with.var%>%select(Pain,Itching)%>%distinct()

#it seems patients who have pain also have itching, so will only evaluate itching
sig.vals<-data.with.var%>%
  spread(key=Itching,value=score)%>%
  group_by(method,cell_type)%>%
  mutate(pval=wilcox.test(`TRUE`,`FALSE`)$p.value)%>%
  select(method,cell_type,pval)%>%distinct()%>%
  ungroup()

#there are some?!
DT::datatable(subset(sig.vals,pval<0.05))

subset(data.with.var,method=='mcp_counter')%>%
  subset(cell_type%in%subset(sig.vals,pval<0.05)$cell_type)%>%
  ggplot()+geom_boxplot(aes(x=cell_type,y=score,fill=Itching))+scale_y_log10()

```

Last thing to check is that if the Itching phenotype is correlated with the growth.

## Next steps

Not sure what else we can do with these. Some ideas:

* Look to see if these cell types are up-regulated in pNFs that become MPNSTs
* Look to see if these cell types are different between pNFs and MPNSTs

