---
title: "10 - pNF progression in Immune Signatures"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()
require(tidyverse)
```

## Synapse tables with data 
For this data we are using the JHU biobank data that we've analyzed previously as part of a larger cohort:
- [Patient Table](https://www.synapse.org/#!Synapse:syn18634454/tables/)
- [RNA-Seq Table](https://www.synapse.org/#!Synapse:syn20812185/tables/) (actually we don't use this, only the transformed data below:)

We also have Immune Data
- [Tumor Deconvolution Table](https://www.synapse.org/#!Synapse:syn20710536/tables/)

Let's see if there are any correlations between immune populations and tumor growth.


```{r,warning=FALSE}
pat.tab<-synapser::synTableQuery("SELECT Patient_ID as individualID,Dermal_Neurofibromas,Plexiform_Neurofibromas,MPNST FROM syn18634454")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))

pat.samps<-synTableQuery('select distinct specimenID,individualID from syn20812185')$asDataFrame()

imm.tab<-synapser::synTableQuery("SELECT * FROM syn20710536")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))%>%left_join(pat.samps,by='specimenID')

tab<-pat.tab%>%left_join(imm.tab,by='individualID')

tab%>%select(individualID,Plexiform_Neurofibromas,MPNST,specimenID,tumorType)%>%distinct()%>%DT::datatable()



```

## Compare patients 

First let's look at pNF immune scores bracketed by whether or not patient has pNFs

```{r, warning=FALSE}
tab%>%subset(tumorType=='Plexiform Neurofibroma')%>%
  subset(method!='xcell')%>%
  ggplot()+geom_boxplot(aes(x=cell_type,y=score,fill=MPNST))+scale_y_log10()+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+facet_grid(.~method)

```

Interesting! Some cell types look different. Let's try the converse.
```{r,warning=FALSE}

tab%>%subset(tumorType=='Plexiform Neurofibroma')%>%
  subset(method!='xcell')%>%
  ggplot()+geom_boxplot(aes(x=cell_type,y=score,fill=`Dermal_Neurofibromas`))+scale_y_log10()+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+facet_grid(.~method)

tab%>%subset(tumorType=='Malignant Peripheral Nerve Sheath Tumor')%>%
  subset(method!='xcell')%>%
  ggplot()+geom_boxplot(aes(x=cell_type,y=score,fill=Plexiform_Neurofibromas))+scale_y_log10()+ theme(axis.text.x = element_text(angle = 90, hjust = 1))+facet_grid(.~method)

```

## Significance tests

For each tumor type, evaluate the significance in differences with/without MPNST development.

```{r,warning=FALSE}

mpnst.res<-tab%>%
  subset(tumorType=='Plexiform Neurofibroma')%>%
  spread(key=MPNST,value=score)%>%
  group_by(method,cell_type)%>%
  mutate(pval=t.test(Present,Absent)$p.value)%>%
  select(method,cell_type,pval)%>%distinct()%>%
  ungroup()%>%
  mutate(correctedP=p.adjust(pval))

mpnst.res%>%arrange(pval)

```

Not very much is significant, but we can try to plot those for further analysis/research.

```{r, warning=FALSE}

toplot<-subset(mpnst.res,pval<0.1)

subset(tab,tumorType=='Plexiform Neurofibroma')%>%
  subset(method%in%toplot$method)%>%
  subset(cell_type%in%toplot$cell_type)%>%
  ggplot(aes(x=cell_type,y=score,col=MPNST))+
    geom_boxplot(outlier.shape=NA)+
    geom_point(position=position_jitterdodge())+
  facet_grid(.~method)+scale_y_log10()
  
```

Because this is only five patients we are limited. What if we look at neurofibromas as well?

## Neurofibromas OR Plexiform Neurofibromas

Here we group neurofibromas with pNFs to see if we get the same result

```{r, warning=FALSE}
nf.res<-tab%>%
  subset(tumorType%in%c('Plexiform Neurofibroma','Neurofibroma'))%>%
  spread(key=MPNST,value=score)%>%
  group_by(method,cell_type)%>%
  mutate(pval=t.test(Present,Absent)$p.value)%>%
  select(method,cell_type,pval)%>%distinct()%>%
  ungroup()%>%
  mutate(correctedP=p.adjust(pval))

nf.res%>%arrange(pval)

toplot<-subset(nf.res,pval<0.1)

for(method in c('cibersort','mcp_counter')){
 
   nplot<-toplot[which(toplot$method==method),]##subset fails here,
  
  p<-tab%>%subset(tumorType%in%c('Plexiform Neurofibroma','Neurofibroma'))%>%
  subset(method==method)%>%
  subset(cell_type%in%nplot$cell_type)%>%
  ggplot(aes(x=cell_type,y=score,col=MPNST))+
    geom_boxplot(outlier.shape=NA)+
    geom_point(position=position_jitterdodge(),aes(shape=tumorType,group=MPNST))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
    ggtitle(method)
  if(method=='cibersort')
    p<-p+scale_y_log10()
  print(p)
}

```

## Next steps
We should look into the information about this, I'm not sure if it's prognostic or not. 
