---
title: "11 - Immune Phenotype by genetic mutation"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(synapser)
synLogin()
require(tidyverse)
```

## Synapse tables with data 
For this data we are using any data for which there are gene variants (cNFs, pNFs, MPNSTs):
- [Exome-Seq variants](https://www.synapse.org/#!Synapse:syn20554939/tables/)
- [WGS Variants](https://www.synapse.org/#!Synapse:syn20551862/tables/) 

We also have Immune Data
- [Tumor Deconvolution Table](https://www.synapse.org/#!Synapse:syn20710536/tables/)

Let's see if there are any correlations between specific variants and immune populations. First we collect all the data and find 'recurrent' mutations. Currently we're focusing on mutations that occur in at least 3 samples. 


```{r,warning=FALSE}

wgs.vars=synTableQuery("SELECT Hugo_Symbol,Protein_position,specimenID,IMPACT,FILTER,ExAC_AF,gnomAD_AF FROM syn20551862")$asDataFrame()
exome.vars=synTableQuery("SELECT Hugo_Symbol,Protein_position,specimenID,IMPACT,FILTER,ExAC_AF,gnomAD_AF FROM syn20554939")$asDataFrame()

all.vars<-rbind(select(wgs.vars,'Hugo_Symbol','Protein_position','specimenID','IMPACT','gnomAD_AF'),
    select(exome.vars,'Hugo_Symbol','Protein_position','specimenID','IMPACT','gnomAD_AF'))%>%
  subset(gnomAD_AF<0.01)

imm.tab<-synapser::synTableQuery("SELECT * FROM syn20710536")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))

samps<-intersect(imm.tab$specimenID,all.vars$specimenID)

tab<-imm.tab%>%subset(specimenID%in%samps)%>%
    left_join(all.vars,by='specimenID')

tab<-subset(tab,!tumorType%in%c('Other','High Grade Glioma','Low Grade Glioma'))

top.genes=tab%>%group_by(tumorType)%>%
  mutate(numSamps=n_distinct(specimenID))%>%
      group_by(tumorType,Hugo_Symbol)%>%
    mutate(numMutated=n_distinct(specimenID))%>%
    ungroup()%>%
  subset(numMutated>1)%>%
      subset(numMutated<(numSamps-1))%>%
  select(tumorType,Hugo_Symbol,numSamps,numMutated)%>%distinct()

gene.count=top.genes%>%group_by(tumorType)%>%mutate(numGenes=n_distinct(Hugo_Symbol))%>%select(tumorType,numGenes)%>%distinct()
DT::datatable(gene.count)
```

Ok, we have some genes and counts that we can look at! 

## Test significance of each gene/immune population

Now we can loop through every tumor type and gene

```{r, warning=FALSE}
red.genes<-c("NF1","SUZ12","CDKN2A","EED")##for testing

vals<-tab%>%#subset(Hugo_Symbol%in%red.genes)%>%
    mutate(mutated=ifelse(is.na(IMPACT),'WT','Mutated'))%>%
  select(cell_type,method,tumorType,score,Hugo_Symbol,specimenID,mutated)%>%
  distinct()%>%
  spread(key=Hugo_Symbol,value='mutated',fill='WT')

counts<-vals%>%
  gather(key=gene,value=status,-c(cell_type,method,tumorType,score,specimenID))%>% 
    select(cell_type,method,tumorType,score,gene,specimenID,status)%>%
    group_by(cell_type,method,tumorType,gene)%>%
    mutate(numVals=n_distinct(status))%>%
    subset(numVals==2)%>%ungroup()

#so now we have only 
with.sig<-counts%>%ungroup()%>%subset(gene%in%top.genes$Hugo_Symbol)%>%
  subset(method!='xcell')%>%
    group_by(method,cell_type,gene)%>%
  mutate(pval=t.test(score~status)$p.value)%>%ungroup()%>%
  group_by(cell_type,method)%>%
  mutate(corP=p.adjust(pval))%>%ungroup()%>%
  select(cell_type,method,gene,pval,corP)%>%distinct()

sig.vals<-subset(with.sig,corP<0.3)

DT::datatable(sig.vals)

```

Interesting! Some genes actually pass p-value correction. What do they look like? Here let's write the messiest possible code to print.

```{r,warning=FALSE}
##i am sooo not proud of this horrible code. apologies!!!
for(method in c('cibersort','mcp_counter')){
  ##first let's reduce the counts and signifiance measures
  rcounts<-subset(counts, method==method)
  nplot<-sig.vals[which(sig.vals$method==method),]##subset fails here,      

  for(ct in unique(nplot$cell_type)){
    tplot<-nplot[which(nplot$cell_type==ct),]
    if(nrow(tplot)==0)
      next
    print(tplot)
    p<-rcounts%>%
    subset(cell_type==ct)%>%
    subset(gene%in%tplot$gene)%>%
    ggplot(aes(x=gene,y=score,col=status))+
    geom_boxplot(outlier.shape=NA)+
    geom_point(position=position_jitterdodge(),aes(group=status))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))+
    ggtitle(paste(method,ct,'predictions'))
    if(method=='cibersort')
      p<-p+scale_y_log10()
    print(p)
  }
}


```



## Next steps
We should look into each individual tumor type, but I was having trouble executing the proper group by's and gave up. Happy for someone else to give it a whirl



```{r, warning=FALSE}

#this is a failed attempt to group by tumor type
#with.sig<-counts%>%ungroup()%>%subset(gene%in%top.genes$Hugo_Symbol)%>%
#    group_by(latent_var,tumorType,gene)%>%
#  mutate(pval=t.test(value~status)$p.value)%>%
#  ungroup()%>%
#  group_by(latent_var)%>%
#  mutate(corP=p.adjust(pval))%>%ungroup()%>%
#  select(latent_var,tumorType,gene,pval,corP)%>%distinct()

#sig.vals<-subset(with.sig,corP<0.05)

#DT::datatable(sig.vals)
```

