---
title: "Tumor type association with latent variable expression in NF tumors"
author: "Robert Allaway"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


Here, we are evaluating the expression of latent variables in NF tumors as they pertain to the type of the tumor. These data were generated as part of developing the braiNFood app. 

# Import packages

First, import packages to process and plot the data. 


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(ggbeeswarm)
library(synapser)
synLogin()
deconv_scores='syn20710536'
dtab<-synapser::synTableQuery(paste('select * from',deconv_scores))$asDataFrame()%>%
  rename(immScore='score')

combined<-subset(dtab,tumorType%in%c('Cutaneous Neurofibroma','Plexiform Neurofibroma','Neurofibroma','Malignant Peripheral Nerve Sheath Tumor'))


```

# Significance testing

We already have the expression of the immune signatures in these tumors. Let's perform a kruskall-wallis test and adjust the resulting p-value based on the expression of a latent variable as a function of the sex of the patient the sample was taken from. 

Then we take the tidy data frame of immune data, group by the variable and tumor type, nest the dataframe based on those groups, and then calculate the p-value for each nested data frame. Nothing is significant, so we plot a boxplot for any latent variable where the p-value is less than 0.3. 

```{r message=FALSE, warning=FALSE}


res_model <- combined%>%
  group_by(method,cell_type) %>% 
  nest() %>% 
  mutate(pval= map(data,function(x){
    kruskal.test(formula = tumorType~ immScore, data = x)$p.value
  }) %>% 
           as.numeric %>% 
           signif(., digits = 3)) %>% 
  mutate(p_adj = p.adjust(pval, n = nrow(.), method = "BH")) %>% 
  filter(pval < 0.3) %>% 
  arrange(-pval) 

#tidy <- plier_loadings_df %>% 
#      tibble::rownames_to_column('gene') %>%
#  gather(lv, loading, -gene) %>% 
#  filter(loading > 0)

 
res_plot <- res_model %>% 
  mutate(title = paste0(method,':',cell_type," p-value = ", pval) %>% 
           str_wrap(., width = 40)) %>% 
  mutate(plots = map2(title, data, function(.x,.y){
      ggplot(data = .y %>% mutate(tumorType = str_wrap(tumorType, width = 15))) +
      geom_boxplot(aes(x = tumorType, y = immScore, fill = tumorType)) +
      geom_beeswarm(aes(x = tumorType, y = immScore)) +
      scale_y_log10()+
      ggtitle(.x) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
  })) #%>% 
 # mutate(plots_loading = map(latent_var, function(x){
#    ggplot(tidy %>% filter(lv == x) %>% top_n(30, loading)) +
#    geom_bar(aes(x=reorder(gene, -loading), y=loading, fill = gene %in% drug_targets$gene), stat = "identity") +
#    scale_fill_discrete(name="Druggable") +
#    theme_bw() +
#    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
#      labs(x = "Gene", y = "LV Loading")
#  }))

DT::datatable(res_model %>% select(-data))
```

# Plots 

### Tumor Specific 

Here are the latent variables where p < 0.2 when grouping by tumor type.

```{r}

#plots <- map(res_plot$plots,function(.x){#res_plot$plots_loading,function(.x,.y){
#    combo_plot <- gridExtra::grid.arrange(.x)#,.y)
#})

res_plot$plots
```

