---
title: "Tumor type association with immune expression in NF tumors"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


Here, we are evaluating the expression of immune cell types in NF tumors as they pertain to the type of the tumor. 
# Import packages

First, import packages to process and plot the data. 


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(stringr)
library(ggbeeswarm)
library(synapser)
synLogin()
deconv_scores='syn20710536'
dtab<-synapser::synTableQuery(paste('select * from',deconv_scores))$asDataFrame()%>%
  rename(immScore='score')

combined<-subset(dtab,tumorType%in%c('Cutaneous Neurofibroma','Plexiform Neurofibroma','Neurofibroma','Malignant Peripheral Nerve Sheath Tumor'))


```

# Significance testing

We already have the expression of the immune signatures in these tumors. Let's perform a kruskall-wallis test and adjust the resulting p-value based on the expression of aimmune cell type.

Then we take the tidy data frame of immune data, group by the cell type and tumor type, nest the dataframe based on those groups, and then calculate the p-value for each nested data frame. After making `tumorType` into a factor I was able to identify significant cell types that passed a BH correction test less than 0.05.

```{r message=FALSE, warning=FALSE}

combined$tumorType<-as.factor(combined$tumorType)
res_model <- combined%>%subset(method!='xcell')%>%
  group_by(method,cell_type) %>% 
  nest() %>% 
  mutate(pval= map(data,function(x){
    kruskal.test(formula = immScore~tumorType, data = x)$p.value
  }) %>% 
           as.numeric %>% 
           signif(., digits = 3)) %>% 
  mutate(p_adj = p.adjust(pval, n = nrow(.), method = "BH")) %>% 
  filter(p_adj < 0.05) %>% 
  arrange(p_adj )

#tidy <- plier_loadings_df %>% 
#      tibble::rownames_to_column('gene') %>%
#  gather(lv, loading, -gene) %>% 
#  filter(loading > 0)

 
res_plot <- res_model %>% 
  mutate(title = paste0(method,':',cell_type," p-value = ", pval) %>% 
           str_wrap(., width = 40)) %>% 
  mutate(plots = map2(title, data, function(.x,.y){
      ggplot(data = .y %>% mutate(tumorType = str_wrap(tumorType, width = 15))) +
      geom_boxplot(aes(x = tumorType, y = immScore, fill = tumorType)) +
      geom_beeswarm(aes(x = tumorType, y = immScore)) +
      scale_y_log10()+
      ggtitle(.x) +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5))
  })) #%>% 
 # mutate(plots_loading = map(latent_var, function(x){
#    ggplot(tidy %>% filter(lv == x) %>% top_n(30, loading)) +
#    geom_bar(aes(x=reorder(gene, -loading), y=loading, fill = gene %in% drug_targets$gene), stat = "identity") +
#    scale_fill_discrete(name="Druggable") +
#    theme_bw() +
#    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
#      labs(x = "Gene", y = "LV Loading")
#  }))

DT::datatable(res_model %>% select(-data))
```

# Plots 

### Tumor Specific 

Here are the immune types  where p < 0.2 when grouping by tumor type.

```{r}

#plots <- map(res_plot$plots,function(.x){#res_plot$plots_loading,function(.x,.y){
#    combo_plot <- gridExtra::grid.arrange(.x)#,.y)
#})

res_plot$plots
```

