---
title: "21-Top Latent Variable - Tumor Deconvolution Correlation"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##compare tumor immune scores to lv expression

library(synapser)
synLogin()
library(tidyverse)

deconv_scores='syn20710536'
```

## Immune predictions correlated with multiplier latent variable expression

First we get the LV multiplier values from the nf-lv-viz repo and deconvolution predictions from Synapse. We filter the LVs based on those in `syn21315356`.

```{r, warning=FALSE}
#get immune predictions
dtab<-synapser::synTableQuery(paste('select * from',deconv_scores))$asDataFrame()%>%
  rename(immScore='score')

##get latent variable scores
mp_res<-synTableQuery("SELECT * FROM syn21046991")$filepath %>% 
  readr::read_csv() %>% 
  filter(!is.na(tumorType),
         !grepl('xenograft', specimenID, ignore.case = T),
         !specimenID %in% c("BI386-004","CW225-001","DW356-002",
                            "JK368-003", "SK436-005"))

#get RF-selected latent variables
lvs<-synTableQuery("select * from syn21318452")$asDataFrame()$LatentVar


dtab<-subset(dtab,method!='xcell')

combined=dtab%>%select(c(cell_type,method,specimenID,immScore))%>%
  left_join(subset(mp_res,latent_var%in%lvs),by='specimenID')
```

Then we can compute the correlation of each cell type measurement with each protein
```{r, warning=FALSE}
corVals=combined%>%group_by(cell_type,latent_var,method)%>%summarize(corVal=cor(immScore,value,method = "spearman"))

DT::datatable(corVals)
```



## Plotting LV correlations

We are interested in particular protein signaling activity that might give rise to specific immune phonetypes.

```{r, warning=FALSE}
##now how do we bracket them?
##plot correlation distributions by cell type and method. 
require(ggplot2)
p<-ggplot(corVals)+geom_boxplot(aes(x=cell_type,y=corVal,fill=method))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ggtitle("Correlation of MultiPLIER Latent Variables with predicted cell type")
print(p)
```

There are some latent variables that show up as highly correlated. By choosing a threshold, we can evaluate what they are in more detail.

```{r, warning=FALSE}
corthresh=0.6

##now filter to the cell types with correlated latent variables (or anticorrelated)
cor_cell_types=subset(corVals,abs(corVal)>corthresh)%>%
  ungroup()%>%
  group_by(latent_var)%>%
  mutate(numTypes=n_distinct(cell_type))%>%
  subset(numTypes>1)%>%
  mutate(cell_types=paste(unique(cell_type),collapse=','))%>%
  ungroup()%>%
  select(latent_var,method,cell_types)%>%unique()
print(paste('we found',nrow(cor_cell_types),'latent variables/methods with some cell types with correlation greater than',corthresh))

DT::datatable(cor_cell_types)
```

## Plots of LVs with correlated immune cell types

Here we iterate through each of the LVs with correlated immune cell types and plot

### Cibersort-correlated cell types

First we plot the cibersort

```{r cor cell type plots}
library(nationalparkcolors)

cells<-subset(corVals,abs(corVal)>corthresh)%>%ungroup()%>%
    select(cell_type)%>%distinct()%>%subset(cell_type!='Monocyte')

val<-park_palette('GeneralGrant',nrow(cells))
names(val)<-cells$cell_type

apply(subset(cor_cell_types),1,function(x){
  lv=x[['latent_var']]
  m=x[['method']]
  ct<-unlist(stsplit(x[['Cell Types']],split=','))
  
  #for each gene and cell type
  lvs=subset(corVals,latent_var==lv)%>%
        subset(abs(corVal)>corthresh)%>%
        subset(cell_type!='Monocyte')%>%
    subset(method==m)%>%arrange(desc(corVal))%>%
      ungroup()

  scores=subset(combined,cell_type%in%ct)%>%
      subset(latent_var==lv)%>%
      subset(method==m)

  p2<- ggplot(scores %>% mutate(latent_var = stringr::str_trunc(latent_var, 30)))+
      geom_point(aes(x=immScore,y=value,
          col=cell_type,shape=tumorType))+theme_bw()+
    scale_color_manual(values=val)+
      ggtitle(paste(m,'predictions of',lv,'correlation >',corthresh))
  if(method=='mcp_counter')
    p2<-p2+scale_x_log10()
  print(p2)
})

```