---
title: "22-Selected LV-Metaviper Cors"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
##compare tumor immune scores to metaviper predictions

require(synapser)
synLogin()
require(tidyverse)

mp_scores='syn21046991'
metaviper_scores='syn21259610'
rf_mp='syn21222255'

```

### Latent Variables active in Random Forests

First we get the metaviper predictions, LV scores, and Random Forest weights from Synapse. We filter for LVs that are selected by the random forest.

```{r, warning=FALSE}
#get immune predictions
dtab<-synapser::synTableQuery(paste('select * from',mp_scores))$asDataFrame()%>%
  subset(isCellLine!='TRUE')

##get metaviper scores
mtab<-synapser::synTableQuery(paste('select * from',metaviper_scores))$asDataFrame()

##get rf loadings
rftab<-synapser::synTableQuery(paste('select * from',rf_mp))$asDataFrame()%>%
  select(LV_Full,`Cutaneous Neurofibroma`,`Neurofibroma`,`Malignant Peripheral Nerve Sheath Tumor`,`Plexiform Neurofibroma`)%>%
  mutate(latent_var=gsub('`','',LV_Full))%>%
  select(-LV_Full)

samps<-intersect(dtab$specimenID,mtab$specimenID)

#get RF-selected latent variables
lvs<-synTableQuery("select * from syn21318452")$asDataFrame()%>%
  rename(latent_var='LatentVar')%>%
  left_join(rftab,by='latent_var')

mp_res<-dtab%>%
  subset(specimenID%in%samps)%>%
  select(latent_var,value,specimenID,diagnosis)%>%
  left_join(rftab,by='latent_var')

combined<-mp_res%>%ungroup()%>%inner_join(mtab,by='specimenID')

#now compute some basic stats
mp_stats<-mp_res%>%
  rowwise()%>%mutate(MaxVal=max(`Cutaneous Neurofibroma`,`Plexiform Neurofibroma`,`Malignant Peripheral Nerve Sheath Tumor`,Neurofibroma))%>%
  rowwise()%>%
  mutate(MeanVal=mean(c(`Cutaneous Neurofibroma`,`Plexiform Neurofibroma`,`Malignant Peripheral Nerve Sheath Tumor`,Neurofibroma)))

DT::datatable(mp_stats)
```


### Plotting protein correlations
With the RF-selected LVs for each random forest prediction, we can plot those metaviper proteins that correlate with them. 


```{r, warning=FALSE}
corVals=combined%>%#subset(latent_var%in%unique(unlist(top10)))%>%
    group_by(latent_var,gene)%>%
  summarize(corVal=cor(value,metaviperscore,use='pairwise.complete.obs'),numSamps=n_distinct(specimenID))

corVals
##let's store this in Synapse
tab<-synBuildTable('Top10 Metaviper Latent-Variable Correlations',parent='syn21046734',corVals)
synStore(tab)

#corVals<-corVals%>%subset(latent_var%in%unique(unlist(top10)))
```


```{r, warning=FALSE}
##now how do we bracket them?
##plot correlation distributions by cell type and method.
require(ggplot2)

##first re-order variables to plot
top.df<-mp_stats%>%rename(All='MeanVal')%>%
  select(-c(specimenID,diagnosis,value,MaxVal))%>%
  gather(key="disease",value="importance",-latent_var)%>%unique()

p<-corVals%>%
              ungroup()%>%
  subset(latent_var%in%unique(unlist(lvs)))%>%
          #    mutate(LatentVariable = stringr::str_trim(as.character(latent_var), 20))%>%
              ggplot()+geom_boxplot(aes(x=latent_var,y=corVal))+ theme(axis.text.x = element_text(angle = 45, hjust = 1))+ggtitle("Correlation of metaviper proteins with lv")
print(p)
```

There are some proteins that show up as highly correlated. By choosing a threshold, we can evaluate what they are in more detail.

These plots represent the top latent variables for a predictor of each tumor type and the proteins that are correlated with them. 

```{r, warning=FALSE}
corthresh=0.75

##now filter to the cell types with correlated proteins
cor_cell_types=subset(corVals,corVal>corthresh)%>%
  subset(latent_var%in%unique(unlist(lvs)))%>%
      ungroup()%>%
  select(latent_var)%>%
  distinct()

print(paste('we found',nrow(cor_cell_types),'lvs with some protein correlation greater than',corthresh))

DT::datatable(cor_cell_types)

apply(cor_cell_types,1,function(x){
  ct=x[['latent_var']]
#  m=x[['method']]

  #for each gene and cell type
  genes=subset(corVals,latent_var==ct)%>%
        subset(corVal>corthresh)%>%
   arrange(desc(corVal))%>%
      ungroup()

    if(nrow(genes)>12){
    new.corthresh=format(genes$corVal[12],digits=3)
    genes=genes[1:12,]
  }else{
    new.corthresh=corthresh
  }

  scores=subset(combined,gene%in%genes$gene)%>%subset(latent_var==ct)

  p2<- ggplot(scores)+
      geom_point(aes(x=value,y=metaviperscore,
          col=gene,shape=tumorType))+
  #  scale_x_log10()+
      ggtitle(paste(ct,'correlation >',new.corthresh,'\n',
        subset(top.df,latent_var==ct)%>%select(disease)%>%unlist()%>%paste(collapse=',')))
  cat(ct)##print out so we can search
  print(p2)
 # ggsave(paste0(m,'predictions of',gsub(" ","",gsub("/","",ct)),'cor',new.corthresh,'.pdf'))
})


```
