---
title: "Latent variable scores correlated with mean MetaViper scores of drugs"
author: "Sara Gosline"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```


Here, we are evaluating the 'druggability' of viper proteins that are well correlated with variables. 
The plots at the end of this are labeled by LVs, but the enrichment is actually based on the list of viper proteins that have a correlation of 0.5 or greater with LV expression across all tumors.

# Import packages

First, import packages to process and plot the data. 

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(synapser)
library(feather)
synLogin()

#get viper scores
viper_res <- synTableQuery('SELECT * FROM syn21259610')$filepath %>% 
readr::read_csv()%>%select(-c(ROW_ID,ROW_VERSION))

#get top LVs
top_lvs<-synTableQuery('SELECT distinct latent_var FROM syn21416832')$asDataFrame()$latent_var

#get multiplier scores
mp_res <-synTableQuery("SELECT * from syn21046991 where tumorType <> 'NA'")$filepath %>%
  readr::read_csv()%>%
  select(latent_var,value,tumorType,specimenID,studyName)%>%
  subset(latent_var%in%top_lvs)

#get drug targets
drug_targets <- feather::read_feather(synGet('syn20700199')$path)
drug_targets <- drug_targets %>% 
  filter(mean_pchembl > 7) %>% 
  mutate(gene= hugo_gene) %>% 
  select(gene, std_name) 
```

# Map drugs and targets to summarize mapping

For each drug we want to compute the mean metaviper score of each of its targets in each of the samples to provide a per-sample analysis of drug efficacy.


```{r message=FALSE, warning=FALSE}

term2viper <- drug_targets %>% 
  mutate(term = std_name) %>% 
  select(term, gene) %>% 
  distinct()%>%
  left_join(viper_res,by='gene')

library(ggplot2)
drugTargs<-term2viper%>%
  group_by(term)%>%
  summarize(numTargs=n_distinct(gene))
 
ggplot(drugTargs)+geom_bar(aes(x=numTargs))
```

Most drugs have only 1 target! Let's filter to those drugs that have at least 3 targets

# Calculate mean drug metaviper values and correlation with LV

```{r}

topDrugs<-subset(drugTargs,numTargs>2)%>%
  select(term)

drugMeans<-term2viper%>%
  subset(term%in%topDrugs$term)%>%
  group_by(term,specimenID)%>%
  summarize(meanViper=mean(metaviperscore))

lv2drug<-drugMeans%>%
  ungroup()%>%
  left_join(mp_res,by='specimenID')

corLvs<-lv2drug%>%
  group_by(latent_var,term)%>%
  summarize(lvDrugCor=cor(value,meanViper,method='spearman'))%>%
  subset(abs(lvDrugCor)>0.65)%>%left_join(drugTargs,by='term')

DT::datatable(corLvs)

```

There are sooo mnay drugs/LVs that are correlated. This is likely due to the high amount of overlap with drugs and targets.

# Filtering for significant drug/lv combos

Since Robert calculated the enrichment of each LV by drug targets in `syn21442304` we can select those with significant (p.adjust<0.01) enrichment that also have drugs that are somewhat recognizable (not `CHEMBL` ids or some form of `Acid`)

```{r}
sigs<-synTableQuery("SELECT * from syn21442304 where \"p.adjust\"<0.01 and ID not like 'CHEMBL%' and ID not like '%Acid'")$asDataFrame()%>%
  select(-c(ROW_ID,ROW_VERSION))

res=sapply(unique(sigs$latent_var),function(lv){
  drugs<-subset(sigs,latent_var==lv)%>%
    select(ID)%>%
    distinct()
  
  terms=subset(corLvs,latent_var==lv)%>%ungroup()%>%select(term)%>%distinct()
  toplot=subset(lv2drug,latent_var==lv)%>%
    subset(term%in%drugs$ID)%>%
    subset(term%in%terms$term)%>%
    rename(`LV Score`='value')
      
  toplot$term<-droplevels(toplot$term)
  p<-ggplot(toplot)+geom_point(aes(x=meanViper,y=`LV Score`,col=term,shape=tumorType))+
    ggtitle(paste(lv,'correlated drugs by Viper score'))
  print(p)
  ggsave(paste(lv,'corDrugs.pdf',sep=''),useDingbats=FALSE)
  p
})
```



```{r}
sessionInfo()
```